Cross Validation Approaches for Penalized Cox Regression

### 2019-05-14 cross-validation.tex
add package citations; change l,exp and log; a few more comments in blue
Edits to 3.1
clean/reply to some previous comments
edit application and discussion section

\newcommand{\ith}{i^\textrm{th}}
\newcommand{\CVE}{\textrm{CVE}}
\newcommand{\logit}{\mbox{logit}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\ones}{\mathbbm{1}}
\newcommand{\indic}[1]{\boldsymbol{1}_{\{ #1 \}}}
\newcommand{\mvec}{\mbox{vec}}
\newcommand{\cov}{\mbox{Cov}}
\newcommand{\eqdist}{\overset{{\cal D}}{=}}
\newcommand{\const}{\mbox{const}}
\newcommand{\iid}{\overset{iid}{\sim}}
\newcommand{\diag}{\mbox{{\bf diag}}}
\newcommand{\Diag}{\mbox{Diag}}

@@

\providecommand{\note}[1]{\textcolor{red}{#1}}

@@

\section*{Abstract}
Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature. Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation. Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere. \note{IS THIS STILL TRUE?  WE SHOULD EDIT THIS SENTENCE: Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.} However, our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability. We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.
\input{main}

>>>>

\newcommand{\ith}{i^\textrm{th}}
\newcommand{\CVE}{\textrm{CVE}}
\newcommand{\MSE}{\textrm{MSE}}
\newcommand{\logit}{\mbox{logit}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\ones}{\mathbbm{1}}
\newcommand{\indic}[1]{\boldsymbol{1}_{\{ #1 \}}}
\newcommand{\mvec}{\mbox{vec}}
\newcommand{\cov}{\mbox{Cov}}
\newcommand{\eqdist}{\overset{{\cal D}}{=}}
\newcommand{\const}{\mbox{const}}
\newcommand{\iid}{\overset{iid}{\sim}}
\newcommand{\diag}{\mbox{{\bf diag}}}
\newcommand{\Diag}{\mbox{Diag}}

@@

\providecommand{\note}[1]{\textcolor{red}{#1}}
\providecommand{\bnote}[1]{\textcolor{blue}{#1}}

@@

\section*{Abstract}
Cross validation is commonly used for selecting tuning parameters in penalized regression, but its use in penalized Cox regression models has received relatively little attention in the literature. Due to its partial likelihood construction, carrying out cross validation for Cox models is not straightforward, and there are several potential approaches for implementation. Here, we propose two new cross-validation methods for Cox regression and compare them to approaches that have been proposed elsewhere. %\note{IS THIS STILL TRUE?  WE SHOULD EDIT THIS SENTENCE: Our simulation studies show that in general, all cross-validation approaches tend to be conservative (i.e., select smaller models than the optimal choice of tuning parameters) for penalized Cox regression models.} 
Our proposed approach of cross-validating the linear predictors seems to offer an attractive balance of performance and numerical stability. We illustrate these advantages using simulated data as well as real data from two genome-wide studies of survival in cancer patients.
\input{main}

### 2019-05-14 main.tex
Fixing number of genes (85 -> 81)
edit application and discussion section

This semiparametric flexibility, however, comes at a cost. The Cox regression approach can estimate relative risks, but without estimating the baseline hazard, it does not make any predictions concerning the absolute failure time for any given individual.  This poses a challenge to assessing the predictive accuracy of a given Cox regression model.  The challenge is particularly relevant for penalized Cox regression \citep{Tibshirani1997,Fan2002}, as the assessment of predictive accuracy via cross-validation is the standard method for selecting the regularization parameter and deciding upon a model.

@@

\par In the Cox model, the hazard function for subject $i$ is given by 
\begin{equation*}
  h_{i}(t) = h_{0}(t) \exp( X_{i}^{T} \beta),
\end{equation*} 
where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for a subject with covariates $X_i$ relative to that baseline.  The estimation of the coefficients is obtained by maximizing a partial likelihood.  Letting $t_i$ denote the time on study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
\begin{equation}
  \label{eq:cox-pl-subj}
  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
\end{equation}
where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
\begin{equation}
  \label{eq:cox-pl-sum}
  L(\beta) =\prod_{i = 1}^{n} L_{i}(\beta).
\end{equation}
Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
\begin{equation}
  \label{eq:obj}
  Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
\end{equation}
where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.

@@

\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together partial likelihood measures from each fold.  An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood. We refer to this method as the \emph{cross-validated linear predictors} approach. To be more specific, we would first obtain $\hat{\beta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:  
\begin{equation}
  \label{eq:cv-lp}
  \hat{\eta}^{cv}_{i} = X_{i}\hat{\beta}^{-k}
\end{equation} 
After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{cv} = ( \hat{\eta}^{cv}_{1},  \hat{\eta}^{cv}_{2} , ...  \hat{\eta}^{cv}_{n})$ for the whole sample is obtained. A partial likelihood can then be built over this set of linear predictors to evaluate the predictive accuracy of the model: 
	\begin{equation} 
	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}^{cv}_{j})}.
	\end{equation}
We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{cv}).$$

@@

Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression. For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples. Given $X$ and a coefficient vector $\beta$, survival times were generated from exponential distribution with rate parameter $\exp(X\beta)$. Censoring status indicators were drawn independently from the binomial distribution. All simulations were implemented in R \citep{R}. We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation.

@@

Our first simulation study concerns the accuracy of the cross-validation approaches introduces in Section~\ref{Sec:methods}.  Accuracy was assesed via several criteria.

@@

\par The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}. With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be
\begin{equation}
	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
\end{equation}
As for the Brier score, KL scores were computed for all CV approaches at the median survival time.  Results were then averaged over the N replications. For both Brier and KL scores, a score represents smaller prediction error, thus indicating better prediction accuracy.

@@

\subsection {Stability}
\label{Sec:stability}
 \bnote{I updated the writing and figure~\ref{Fig:stability} for this section}

@@

The following simulation study was designed to illustrate this issue. We generated data sets with $n = 120$ observations. The censoring rate is fixed exactly at $50\%$, thus there are 60 events observed among all observations. For each data set, we conducted the balanced fold assiggment with number of folds set to be 10, 12, 15, 20, 30 and 60. When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there were one event in each fold. After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}. We recorded the model that minimizes the cross-validated error for each approach and computed its squared error loss. We replicated the whole process for $N = 200$ time. In the end, instead of looking at the mean as in Section~\ref{Sec:accuracy}, we computed the standard deviation of the squared error loss for each approach.

@@

\bnote{I edited this section and the figure caption}

@@

\subsection {Cross-Validated C Index}
\label{Sec:CIndex}
\bnote{unified the naming of C index in this manuscript. Use Harrell's C Index or C Index instead of other alternatives. This is the default concordance statistic that the survival package would return.}

@@

In this section, we demonstrate the usage and performance of the aforementioned cross validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer. \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients.
They gathered expression profiles of 22,283 genes for 442 non-small-cell lung cancer (NSCLC) patients along with some additional clinical and demographic data. About $50\%$ of the patients died during the course of the study (236 events), with the rest censored. The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival.

\begin{table}[h]
\centering
\begin{tabular}{ccccccccccc}
\toprule
 & \multicolumn{2}{c}{$\alpha = 0.3$ } & \multicolumn{2}{c}{$\alpha = 0.5$} & \multicolumn{2}{c}{$\alpha = 0.7$} & \multicolumn{2}{c}{$\alpha = 0.9$} & \multicolumn{2}{c}{$\alpha = 1$} \\
 &      $\lambda$      &   CVE         &   $\lambda$        &     CVE       &      $\lambda$      &     CVE       &    $\lambda$        &    CVE        &     $\lambda$      &    CVE       \\

\hline\\

 V$\&$VH  & 0.363 &  296.32  &  0.228   &  296.35  & 0.163 &  296.34  &  0.133  &   296.31  &  0.119  &  296.30  \\
 Standard & 0.301 & 145.17  & 0.218  &  145.38 & 0.163  & 145.37 &  0.127 &  145.35 & 0.114   & 145.34 \\
 Linear Pred & 0.288 & 10.661  & 0.181  & 10.673 & 0.163 & 10.673    & 0.127  & 10.672  &    0.114 &  10.672  \\
 Dev Resid & 0.316 & 571.40 & 0.228  &  571.96 & 0.163 & 571.95  & 0.133  &   571.79 &  0.119  &  571.70  \\
\bottomrule
\end{tabular}
\caption{\bnote{Not sure whether to include this table or not}}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{lcccccc}
  \toprule
  & $\lambda$ & log($\lambda$) & $\#$ of Genes Selected & CV C-Index & Brier Score \\ 
\cline{2-6} \\ 
V $\&$ VH& 0.363 & -1.013 &  & 0.639 & 0.243\\
standard & 0.301 & -1.200 &  & 0.647& 0.232\\
Linear Pred &0.288 & -1.246 &  & 0.650&  0.229\\
Dev Resid & 0.316 & -1.153 &  & 0.645& 0.235\\ 
\bottomrule
\end{tabular}
\caption{\label{tb:shedden}Results of elasticnet model with $\alpha = 0.3$ on the Lung Cancer Data Set. The linear Predictor approach picked the smallest $\lambda$. The corresponding model yields the highest CV C-Index and smallest Brier Score evaluated at median survvial time \bnote{I got different number of genes for LP approach...}}
\end{table}

For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}. Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression. For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error. The standard approach, linear predictor approach and deviance approach all reach the lowest cross-validation error at $\alpha = 0.3$.

We report the predictive performance for all four models at $\alpha = 0.3$ and their corresponding $\lambda$s in Table~\ref{tb:shedden}. For each model, we computed the cross-validated C-Index and Brier Scores at predicting the median survival time. How the CV methods perform on real data reflects what is shown in simulation studies. The standard approach and the linear predictor approach, selecting smaller $\lambda$ values, were the more liberal, while Verweij and Van Houwelingen's CVE approaches were both more conversative. With the highest C Index and lowest Brier Score, the linear predictor approach has best predictive performance among all four methods. The same patterns are repeated when $\alpha$ is at different values.

We compared the CVE along the solution path for different methods. In Figure~\ref{fig:CVE}, the Cross Validated Error is rescaled and plotted for all four methods. A $\lambda$ is selected when CVE curve reaches its lower point. The blue line and green line, which represent the linear predictor approach and standard approah, have more curvature near theirs lowest point, making it is easier to pick out the minimum point.

\begin{figure}[h]
  \centering
  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.jpg}
  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$. The horizonal axis is $\log(\lambda)$. The vertical axis is the cross-validated error rescaled to a range from 0 to 1. After the rescaling, the Cross Validated Error of four different methods can be plotted together for comparison. }
\end{figure}

For the elastic net model with $\lam$ and $\alpha$ selected by the cross-validated linear predictor approach, 81 genes were selected.  Of those, 78 were well-annotated.  We carried out an enrichment analysis of these 78 genes using ToppGene \citep{Chen2009c}; the five most significantly enriched molecular functions are listed in Table~\ref{Tab:go}.

@@

\bnote{(END)}
\section{Discussion}

\par Four methods of conducting cross-validation for Cox Regression were compared through simulation studies in terms of their predictive accuracy. All cross-validation methods tend to select the more conservative model: with greater value of $\lambda$, and fewer variables selected into the model. Despite of the stability issue, the standard approach is the most liberal one across all scenarios and yield the least MSE. The proposed linear predictor approach performs closely to the standard approach. Under LOOCV, the linear predictor approach out-performs the standardstandard approach. Both Verweij and Van Houwelingen's approach and the proposed Deviance Residual approach perform conservatively, hence not recommended for use in practice. Results from real data application reflects results from the simulation studies.

@@

\section{Appendix}

\begin{figure}[h]
	\centering
	\begin{minipage}[t]{10cm}
	\begin{algorithm}[H]
	\caption{Standard Method}
	\par \For{k in 1 to K } 
	{
		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
	}

	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
 	\end{algorithm}

	\begin{algorithm}[H]
  	 \caption{Verweij and Van Houwelingen's Method}
	\par \For{k in 1 to K } 
	{
		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
	}

	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
	\end{algorithm}

	\begin{algorithm}[H]
    	\caption{Cross-Validated Linear Predictors}
	\par \For{k in 1 to K } 
	{
		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
	}

	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
	\end{algorithm}

	\begin{algorithm}[H]
   	\caption{Sums of Squared Deviance Residuals}
	\par \For{k in 1 to K } 
	{
		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
	} 
	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
	\end{algorithm}
	\end{minipage}
	\caption{Pseudo Code for Cross-validation Procedures}
	\end{figure}

>>>>

This semiparametric flexibility, however, comes at a cost. The Cox regression model can estimate relative risks, but without estimating the baseline hazard, it does not make any predictions concerning the absolute failure time for any given individual.  This poses a challenge to assessing the predictive accuracy of a given Cox regression model.  The challenge is particularly relevant for penalized Cox regression \citep{Tibshirani1997,Fan2002}, as the assessment of predictive accuracy via cross-validation is the standard method for selecting the regularization parameter and deciding upon a model.

@@

\par In the Cox model, the hazard function for subject $i$ is given by 
\begin{equation*}
  h_{i}(t) = h_{0}(t) \exp( X_{i}^{T} \beta),
\end{equation*} 
where $h_{0}$ is the baseline hazard and $e^{X_i^{T} \beta}$ is the risk for a subject with covariates $X_i$ relative to that baseline.  The estimation of the coefficients is obtained by maximizing a partial likelihood.  Letting $t_i$ denote the time of being in the study for subject $i$ and $\delta_{i}$ indicate whether or not an event is observed for subject $i$, each subject's individual contribution to that partial likelihood is
\begin{equation}
  \label{eq:cox-pl-subj}
  L_{i}(\beta) = \left \{\frac{\exp ( X_{i}^{T} \beta)}{\sum_{ k \in R(t_{i})}\exp ( X_{k}^{T} \beta)}\right \}^{\delta_{i}},
\end{equation}
where $R(t_{i})$ denotes the set of subjects at risk at time $t_{i}$.  The partial likelihood for the entire sample of $n$ subjects is then given by:
\begin{equation}
  \label{eq:cox-pl-sum}
  L(\beta) =\prod_{i = 1}^{n} L_{i}(\beta).
\end{equation}
Cox regression can be extended by introducing a penalty into the partial likelihood.  In penalized Cox regression, $\beta$ coefficient estimates are obtained by minimizing the objective function
\begin{equation}
  \label{eq:obj}
  Q(\beta) = - \frac{1}{n} \log L(\beta) + P_{\lambda}(\beta),
\end{equation}
where $P_{\lambda}(\beta)$ is a penalty function that depends on a regularization parameter $\lam$.

@@

\par Both of the methods in Section~\ref{Sec:cox-cv-existing} consist of adding together log partial likelihood measures from each fold.  An alternative method is to obtain linear predictors for each fold, then combine these linear predictors to calculate a partial likelihood. We refer to this method as the \emph{cross-validated linear predictors} approach. To be more specific, we would first obtain $\hat{\beta}^{-k}$ from training set $T_{k}$; then the cross-validated linear predictors can be calculated for each observation $i$ from the test set $D_k$:  
\begin{equation}
  \label{eq:cv-lp}
  \hat{\eta}^{cv}_{i} = X_{i}\hat{\beta}^{-k}
\end{equation} 
After repeating this for all K folds, a complete set of cross-validated linear predictors $\hat{\eta}^{cv} = ( \hat{\eta}^{cv}_{1},  \hat{\eta}^{cv}_{2} , ...  \hat{\eta}^{cv}_{n})$ for the whole sample is obtained. A partial likelihood can then be built over this set of linear predictors to evaluate the predictive accuracy of the model: 
	\begin{equation} 
	L(\hat{\eta}^{cv}) = \prod_{i=1}^{n} \frac{\exp (\hat{\eta}^{cv}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}^{cv}_{j})}.
	\end{equation}
We define the cross-validation error evaluated with cross-validated linear predictors to be $$\CVE = - 2 \log L(\hat{\eta}^{cv}).$$

@@

Simulation studies were conducted to compare how the methods described in Section~\ref{Sec:methods} behave relative to each other in selecting the tuning parameter $\lambda$ and the corresponding model for LASSO penalized regression. For these simulation studies, the covariate matrix $X$, consisting of $n$ observations and $p$ covariates, was generated from independent $\text{Normal}(0, 1)$ samples. Given $X$ and a coefficient vector $\beta$, survival times were generated from exponential distribution with rate parameter $\exp(X\beta)$. Censoring status indicators were drawn independently from the binomial distribution. All simulations were implemented in R \citep{R}.

@@

Our first simulation study concerns the accuracy of the cross-validation approaches introduces in Section~\ref{Sec:methods}. We varied the sample size, dimension, censoring percentage and signal strength of the simulated data sets to examine how these factors affect the performance of cross-validation. Accuracy was assesed via several criteria.

@@

\par The third measure we used is the Kullback-Leibler score, which measures the log-likelihood of the prediction model evaluated at the observations \citep{VanHouwelingen2011}. With the same notation defined in the previous paragraph, the Kullback-Leibler score is defined to be
\begin{equation}
	\text{KL}(y, \hat{S}(t_0|\lambda,x)) = -\left\{ y\log(\hat{S}(t_0|\lambda,x) + (1 - y)\log(1 - \hat{S}(t_0|\lambda,x)) \right\}.
\end{equation}
As for the Brier score, KL scores were computed for all CV approaches at the median survival time.  Results were then averaged over the N replications. For both Brier and KL scores, a smaller score represents smaller prediction error, thus indicating better prediction accuracy.

@@

\subsection {Stability}
\label{Sec:stability}

@@

The following simulation study was designed to illustrate this issue. We generated data sets with $n = 120$ observations. The censoring rate is fixed exactly at $50\%$, thus there are 60 events observed among all observations. For each data set, we conducted the balanced fold assignment with number of folds set to be 10, 12, 15, 20, 30 and 60. When the number of folds was set to be 10, there were exactly 6 events in each fold; when the number of folds was set to be 60, there was exactly one event in each fold. After each fold assignment, we computed the cross-validated error using all four approaches described in Section~\ref{Sec:accuracy}. We recorded the model that minimizes the cross-validated error for each approach and computed its squared error loss. We replicated the whole process for $N = 200$ time. In the end, we computed the standard deviation of the squared error loss for each approach.

@@

@@

\subsection {Cross-Validated C Index}
\label{Sec:CIndex}

@@

In this section, we demonstrate the usage and performance of the aforementioned cross-validation methods when fitting penalized regression models on data from a study of gene expression and survival in patients with lung cancer. \citet{shedden2008gene} conducted a large, retrospective study to validate prognostic gene expression signatures as predictors of overall survival in lung cancer patients. They gathered expression profiles of 22,283 genes for 442 non-small-cell lung cancer (NSCLC) patients along with some additional clinical and demographic data. About $50\%$ of the patients died during the course of the study (236 events), with the rest censored.

For this analysis, we fit elastic net models to the data in addition to lasso models, as these models often offer improvement over the lasso when many correlated genes are involved \citep{Zou2005}. The clinical variables stage, age and gender were included in the model without penalization, as it is reasonable to assume that all three have an effect on survival. Besides the regularization parameter $\lam$, the elastic net model has an extra tuning parameter $\alpha$, where $\alpha=1$ corresponds to lasso and $\alpha=0$ corresponds to ridge regression. For each $\alpha = 0.3, 0.5, 0.7, 0.9 \text{ and } 1$, we selected a regularization parameter $\lam$ that has the lowest cross-validation error for each of the four cross-validation methods described in Section~\ref{Sec:methods}. We reported the cross-validated C-Index and Brier Scores of thoses models in Table~\ref{tb:shedden}. With the highest C Index and lowest Brier Score, the model selected by the linear predictor approach at $\alpha = 0.3$ has best predictive performance among all of them.

\begin{table}[ht]
\centering
\caption{\label{tb:shedden} The C Index and Brier Score of all elasticnet models selected by cross-validation methods }
\begin{tabular}{lccccccccccc}
\hline
  & \multicolumn{5}{c}{Cross-validated C Index} & \multicolumn{5}{c}{Brier Score} \\ 
  & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ &  
 & $\alpha = 0.3$ & $\alpha = 0.5$ & $\alpha = 0.7$ & $\alpha = 0.9$ & $\alpha = 1$ \\ \cline{2-6} \cline{8-12} 
V $\&$ VH      & 0.639 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.243 & 0.244 & 0.244 & 0.245 & 0.245 \\
Standard       & 0.647 & 0.639 & 0.638 & 0.638 & 0.639 &  & 0.233 & 0.243 & 0.244 & 0.244 & 0.244 \\
Linear Pred    & \textbf{0.650} & 0.647 & 0.638 & 0.638 & 0.639 &  & \textbf{0.229} & 0.232 & 0.244 & 0.244 & 0.244 \\
Deviance Resid & 0.645 & 0.637 & 0.638 & 0.637 & 0.637 &  & 0.235 & 0.244 & 0.244 & 0.245 & 0.245 \\ \hline
\end{tabular}
\end{table}

For each fixed $\alpha$, we compared the cross-validated errors(CVE) from the four cross-validation methods along the solution path. At $\alpha = 0.3$, CVEs are rescaled and plotted for all four met  hods at the same spectrum of $\lambda$ values In Figure~\ref{fig:CVE}. A $\lambda$ is selected when a CVE curve reaches its lower point. The blue line and the green line, which represent the CVE curves of linear predictor approach and standard approah, have more curvature near theirs lowest point, making it easier to pick out the minimum points. The selected $\lambda$ values, represented by the vertical dashed lines, and their corresponding models, reflected what is shown in simulation studies. The linear predictor approach, selecting smallest $\lambda$ values, was most liberal and selected 85 genes. While the standard approach and deviance residual approach selected 68 genes and 60 genes respectively, Verweij and Van Houwelingen's approach only selected 34 genes and performed most conservatively on this data set. Even though we only illustrated the models at $\alpha = 0.3$, same patterns are repeated when $\alpha$ is fixed at other values.

\begin{figure}[h]
  \centering
  \includegraphics[height= 8cm ]{./manuscript_figure/shedden.png}
  \caption{\label{fig:CVE}Rescaled Cross Validated Errors at $\alpha = 0.3$. The horizonal axis is $\log(\lambda)$. The vertical axis is the cross-validated error rescaled to a range from 0 to 1. After the rescaling, the Cross Validated Error of four different methods can be plotted together for comparison. The vertical dashed lines represent the $\lambda$ selected by a specific method. The linear predictor approach selected the smallest $\lambda$ and Verweij and Van Houwelingen's approach selected the greatest $\lambda$}
\end{figure}

We carried out further data analysis based on the elastic net model with $\lam = 0.650$ and $\alpha = 0.3$, selected by the cross-validated linear predictor approach. Of the 85 genes that were selected by this model, 78 were well-annotated.  We carried out an enrichment analysis of these 78 genes using ToppGene \citep{Chen2009c}; the five most significantly enriched molecular functions are listed in Table~\ref{Tab:go}.

@@

\section{Discussion}

Conducting cross-validation is a crutial step in fitting penalized Cox regression model. In this paper, we proposed two intuitive ways to conduct cross-validation for Cox Regression: cross-validated linear predictor and cross-validated deviance residuals. We conducted a series of simulation experiments to study and compare their performance with two existing approaches, the standard cross-validation and Verweij and Van Houwelingen's cross-validated partial likelihood, in selecting tuning parameters when fitting penalized regression models.

Despite the fact Verweij and Van Houwelingen's cross-validated partial likelihood has been widely implemented, its predictive performance seems to suffer, particularly when the dimension of the data grows and signals become weak. As is illustrated in both simulation studies and real data example, Verweij and Van Houwelingen's approach tends to perform conversatively, selecting fewer genes than all other models. In the lung cancer data example, the number of genes selected by other methods almost doubled the number of genes selected by Verweij and Van Houwelingen's approach. From an application standpoint, using Verweij and Van Houwelingen's approach in high-dimensional datasets may lead to more false negatives and miss important true findings.

While the standard approach was the most liberal one across all simulation scenarios and yielded the best predictive performance, this approach limited by its stability issues which are carefully discussed in Section~\ref{Sec:stability}. The cross-validated linear predictor approach performs closely to the standard approach and demostrates an attractive balance of predictive performance and numerical stability.  We would recommend using the proposed linear predictor approach in practice for selecting regularization parameters in penalized regression.

@@

%\section{Appendix}

%\begin{figure}[h]
%	\centering
%	\begin{minipage}[t]{10cm}
%	\begin{algorithm}[H]
%	\caption{Standard Method}
%	\par \For{k in 1 to K } 
%	{
%		 Fit Cox Regression on $T_{k}$: Obtain model estimates $\hat{\beta}^{-k}$\;
%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D_{k}$: $l^{k}(\hat{\beta}^{-k})$\;		
%	} 
%	
%	\par $ CVE_{st} = - \sum_{k = 1}^{K} l^{k}(\hat{\beta}^{-k})$
% 	\end{algorithm}
%
%	\begin{algorithm}[H]
%  	 \caption{Verweij and Van Houwelingen's Method}
%	\par \For{k in 1 to K } 
%	{
%		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $T_{k}$: $l^{-k}(\hat{\beta}^{-k})$\;
%		 Evaluate the partial likelihood of $\hat{\beta}^{-k}$ on $D$: $l(\hat{\beta}^{-k})$\;
%		 Take the difference: $ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) $\;
%	} 
%
%	\par $ CVE_{vvh} = - \sum_{k = 1}^{K} \left\{ l(\hat{\beta}^{-k}) - l^{-k}(\hat{\beta}^{-k}) \right\}$
%	\end{algorithm}
%
%	\begin{algorithm}[H]
%    	\caption{Cross-Validated Linear Predictors}
%	\par \For{k in 1 to K } 
%	{
%		 Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
%	} 
%
%	\par $ CVE_{lp} = - l(\hat{\eta}^{-})$
%	\end{algorithm}
%	
%	\begin{algorithm}[H]
%   	\caption{Sums of Squared Deviance Residuals}
%	\par \For{k in 1 to K } 
%	{
%		Fit Cox Regression on $T_{k}$:  Obtain model estimates $\hat{\beta}^{-k}$\;
%		$\forall$ $ i$ subject in $D_k$: $\hat{\eta}_{i}^{-} = X_{i}\hat{\beta}^{-k}$\;
%	} 
%	Compute baseline hazard $\hat{\Lambda}_{0}$ based on $\hat{\eta}^{-}$\;
%	$\forall$ $ i$ subject $\in \left\{1,2,...,n\right\}$, compute:
%		\par\hspace{0.5cm} $\hat{M_{i}} = \delta_{i} -\hat{\Lambda}_{0}(t_{i})e^{\hat{\eta}^{-}_{i}}$;
%		\par\hspace{0.5cm} $d_{i} = sgn(\hat{M}_{i})(-2(\hat{M}_{i} + \delta_{i}log(\delta_{i} - \hat{M}_{i})))^{1/2}$;
%	\par $ CVE_{dr} = \sum_{i = 1}^{n} {d}_{i}^2$ \;
%	\end{algorithm}
%	\end{minipage}
%	\caption{Pseudo Code for Cross-validation Procedures}
%	\end{figure}
